---
title: "Demonstration of Latent-TL Pipeline Using Synthetic Data"
output: html_document
date: "2024-03-25"
---

```{r setup, include=FALSE}
library(poLCA)
library(combinat)
library(glmnet)
```
### 0. Load functions
```{r}
source("load_functions_demo.R")
set.seed(2023)
```

### 1. Set parameters of the true model.
#### We generate the data with $C=4$ subpopulations, $q=20$ manifest variables, and $p=2$ covariates in the causal models. Data of $K=8$ hospitals were generated with a sample size of 1000 in each hospital.

```{r}
C = 4 # Number of subpopulations
q = 20 # Number of manifest variables (chronic conditions)
p = 2 # Number of covariates
K = 8 # Number of hospitals, the first site is the target hospital
n.vec = c(rep(1000,K)) # Sample sizes of hospitals

## Parameters of the MLCA model
Lambda <-  matrix(c(rep(1,C)/C,
                        c(0.2,0.2,0.2,0.4),
                        c(0.1,0.2,0.3,0.4),
                        c(0.4,0.3,0.2,0.1),
                        c(0.3,0.2,0.4,0.1),
                        c(0.2,0.1,0.3,0.4),
                        c(0.5,0.2,0.2,0.1),
                        c(0.2,0.1,0.2,0.5)),nrow=8, ncol=C,byrow = TRUE) # Hospital mixing proportions
Pi <- matrix(runif(q*C,0.05,0.95), q, C) # Parameters of MLCA model

## Parameters of causal models
mu <- matrix(runif(2*K,0,2), K, 2) # Means of covariate distributions, which are different across sites
alpha <- c(0,0.2,0.5) # Parameters of PS model
beta <- c(0.5, 1.5, 1.5) # Coefficient of intercept and covariates in the outcome model
tau <- -1.5  # Treatment effect in the outcome model
```

### 2. Simulate data
```{r}
SimSiteData <- function(k){
  n <- n.vec[k]
  class <- apply(rmultinom(n,1,Lambda[k,]),2, which.max)
  X.lca <- t(matrix(rbinom(q*n, 1, Pi[,class]), q, n))
  X1 <- rnorm(n,mean = mu[k,1], sd = 1)
  X2 <- rnorm(n,mean = mu[k,1], sd = 1)
  trtprobs <- plogis(alpha%*%rbind(1,X1,X2))
  Z <- rbinom(n, 1, trtprobs)
  outprobs <- plogis(c(beta,tau)%*%rbind(1,X1,X2,Z))
  Y <- rbinom(n, 1, outprobs)
  
  return(data.frame(S = k, class = class, X.lca = X.lca, X1 = X1, X2 = X2, Z = Z, Y = Y))
}
KSiteData0 <- lapply(1:K,SimSiteData)
KSiteData <- do.call(rbind,KSiteData0)
```

### 3. Perform MLCA to identify subpopulations
```{r}
K <- max(KSiteData$S)
n.vec <- sapply(1:K, function(x) sum(KSiteData$S==x))
X.lca <- KSiteData[,3:(q+2)]
X.lca <- data.frame(X.lca)+1
site <- c(unlist(sapply(1:K, function(k) c(rep(k,n.vec[k])))))

fmla1 <- as.formula(paste('cbind(',paste(colnames(X.lca),collapse = ','),')~site'))
dat.LCA <- cbind(X.lca,site)
set.seed(2)
LCA_train <- poLCA(formula=fmla1, data=dat.LCA,nclass=C,nrep=5, verbose = FALSE)
w.allSite0 <- LCA_train$posterior
```

#### Compare the estimated class membership with true membership: 
```{r}
class.true <- KSiteData$class
class.est0 <- apply(w.allSite0,1,which.max)

match.res <- match_label(class.est0,class.true)
w.allSite <- w.allSite0%*%match.res$Qperm
class.est <- apply(w.allSite0,1,which.max)
KSiteData$class.est <- class.est

match.res$prop_match
```


### 4. Estimation of Target Average Treatment Effect (TATE)
#### Consider subpopulation 1 of hospital 1 as an example. Set up the outcome model, propensity score model, and a model for population standardization as follows:
```{r}
outcome_formula = "Y ~ X1 + X2"
propensity_formula = "Z ~ X1 + X2"
density_ratio_formula = "S ~ X1 + X2"
```

#### 4.1 Benchmark 
#### Estimate the TATE using only data from the target hospital (hospital 1)
```{r}
est.raw <- calATE(k=1, c=1, KSiteData)
est.raw
```

#### 4.2 Transfer learning approach with population standardization and adaptive aggregation
#### Estimate the TATE via transfer learning approach which uses data from hospital 1 as target data and all other hospitals (hospitals 2 to 8) as source data sets.
```{r}
est.TL <- calTATE.agg.CI(k.target=1, source.site=c(2:K), c=1, KSiteData, lambda.lower.bound = NULL)
est.TL$CI[3:4,]
```

